% Copyright (c) 2012 Richard Simard, Pierre L'Ecuyer, Université de Montréal.
% 
% This file is part of Lattice Builder.
% 
% Lattice Builder is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
% 
% Lattice Builder is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
% 
% You should have received a copy of the GNU General Public License
% along with Lattice Builder.  If not, see <http://www.gnu.org/licenses/>.

\defmodule {Base}

This class represents a basis for a lattice.
To compute the length of vectors, one may use either 
the $L_1$, the $L_2$ or the $L_\infty$ norms. If one uses
the $L_2$ norm and if \texttt{BScal} and \texttt{NScal} 
are of type \texttt{double}, then the norm could
overflow if the  components of the  basis vectors are larger than
$\approx 2^{500}$. In that case, \texttt{NScal} may be chosen as
\texttt{RR}.  If one uses
the $L_1$ or the $L_\infty$ norms, \texttt{NScal} may be chosen
the same as  \texttt{BScal}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule
\code\hide
#ifndef BASE_H
#define BASE_H
\endhide
#include "latcommon/Const.h"
#include "latcommon/Types.h"
#include <string>


namespace LatCommon {
 
class Base : public BMat {
public:

   Base (int d, int maxDim, NormType norm = L2NORM);
\endcode
\tabb
Constructor. Builds a basis of actual dimension $d$, maximum dimension
 \texttt{maxDim} and with norm \texttt{norm}.
\endtabb
\code

   Base (const Base &);
\endcode
\tabb
Copy constructor.
\endtabb
\code

   ~Base ();
\endcode
\tabb
Destructor.
\endtabb
\code

   Base & operator= (const Base &);
\endcode
\tabb
Assignment operator.
\endtabb
\code
   
   void kill ();
\endcode
\tabb
Cleans and releases all memory used by the basis.
\endtabb
\code

   void swap (Base & b);
\endcode
\tabb
Swaps this basis with the basis \texttt{b}.
\endtabb
\code

   void permute (int i, int j);
\endcode
\tabb
Exchanges vectors $i$ and $j$ in the basis.
\endtabb
\code

   int getDim () const \hide { return m_dim; } \endhide
\endcode
\tabb
Returns the actual dimension of the basis.
\endtabb
\code
   
   void setDim (int d);
\endcode
\tabb
Resets the actual dimension of the basis to \texttt{d}, but does 
not reserve any new memory.
\endtabb
\code

   int getMaxDim () const \hide { return m_maxDim; } \endhide
\endcode
\tabb
Returns the maximal dimension of the basis.
\endtabb
\code

   NormType getNorm () const \hide { return m_norm; } \endhide
\endcode
\tabb
Returns the norm used by the basis.
\endtabb
\code

   void setNorm (NormType norm);
\endcode
\tabb
Sets the norm used by the basis.
\endtabb
\code

   bool isNegativeNorm (int i) const \hide { return m_negFlag[i]; } \endhide
\endcode
\tabb
Returns \texttt{true} if the $i$-th vector's norm is not updated and
 erroneous. Returns \texttt{false} otherwise.
\endtabb
\code

   void setNegativeNorm (bool flag);
\endcode
\tabb
Sets the dirty flag of the norms to \texttt{flag}.
\endtabb
\code

   void setNegativeNorm (bool flag, int j) \hide { m_negFlag[j] = flag; }\endhide
\endcode
\tabb
Sets the negative flag for the $j$-th vector to \texttt{flag}.
\endtabb
\code

   NScal getVecNorm (int i) const \hide { return m_vecNorm[i]; } \endhide
\endcode
\tabb
Returns the $i$-th vector's norm.
\endtabb
\code

   void setVecNorm (NScal & value, int i);
\endcode
\tabb
Sets the $i$-th vector's norm to \texttt{value}. The negative flag for this
vector is set to false and no data integrity is made.
\endtabb
\code

   void updateVecNorm (); 
\endcode
\tabb
Recalculates the norm of each vector in the basis.
\endtabb
\code

   void updateVecNorm (int d); 
\endcode
\tabb
Same as above, except that the recalculation begins at dimension \texttt{d} + 1.
\endtabb
\code

   void updateScalL2Norm (int d); 
\endcode
\tabb
Updates the norm of vector at dimension \texttt{d} using the \texttt{L2NORM}.
\endtabb
\code

   void updateScalL2Norm (int d1, int d2); 
\endcode
\tabb
Updates the norm of all basis vectors from dimensions \texttt{d1} to
\texttt{d2} (inclusive) using the \texttt{L2NORM}.
\endtabb
\code

   std::string toString () const;
\endcode
\tabb
Writes the basis in a string and returns it.
\endtabb
\code

   std::string toString (int i) const;
\endcode
\tabb
Writes the $i$-th basis vector in a string and returns it.
\endtabb
\code
   
   void write () const;
\endcode
\tabb
Writes the basis on standard output.
\endtabb
\code
 
   void write (int i) const;
\endcode
\tabb
Writes $i$-th basis vector on standard output.
\endtabb
\code


protected:

   int m_dim;
\endcode
\tabb
Actual dimension of the basis.
\endtabb
\code

   int m_maxDim;
\endcode
\tabb
Maximum dimension of the basis.
\endtabb
\code

   NormType m_norm;
\endcode
\tabb
Norm used to calculate the norm of the vectors.
\endtabb
\code

   NVect m_vecNorm;
\endcode
\tabb
The norm of each vector in the basis.
\endtabb
\code

   bool* m_negFlag;     
\endcode
\tabb
Indicates whether a vector norm must be recalculated or not.
\endtabb
\code
};

}
\hide
#endif
\endhide
\endcode
