% Copyright (c) 2012 Richard Simard, Pierre L'Ecuyer, Université de Montréal.
% 
% This file is part of Lattice Builder.
% 
% Lattice Builder is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
% 
% Lattice Builder is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
% 
% You should have received a copy of the GNU General Public License
% along with Lattice Builder.  If not, see <http://www.gnu.org/licenses/>.

\defmodule {Normalizer}

Classes which inherit from this base class are used in implementing bounds
on the length of the shortest nonzero vector in a lattice \cite{mCON99a}.
These bounds are used to normalize the length of the shortest vectors. 
Tight lower bounds are available for all dimensions for many important cases.
In most cases, the ${\cal L}_2$ norm is used to compute the length of vectors.

For some figures of merit, no useful bounds are known to normalize the length
of the shortest vector. In these cases, this base class will be used as 
normalizer since it simply sets all normalization constants to 1. This is
necessary because the tests compare the normalized values of the merit 
when searching for good lattices.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule
\code\hide
/* Normalizer.h for ISO C++ */
#ifndef NORMALIZER_H
#define NORMALIZER_H
\endhide 
#include "latcommon/Types.h"
#include "latcommon/UtilLC.h"
#include "latcommon/Const.h"
#include <string>


namespace LatCommon {
          
class Normalizer {
public:

   static const int MAX_DIM = 48;

   Normalizer (const MScal & m, int k, int t, std::string Name,
               NormType norm = L2NORM, double beta = 1);
\endcode
 \tabb  Constructor for the bounds. Deals with lattices of rank $k$, having
   $m$ points per unit volume, in all dimensions  $\le t$. \texttt{Name}
   is the name of the Normalizer.
   The bias factor \texttt{beta} $= \beta$ gives more weight to some of the
   dimensions: taking $\beta < 1$ inflates the figure of merit by
   $(1/\beta)^t$, thus weakening the requirements for large $t$ in a 
   worst-case figure of merit.  % such as (\ref{eq:M}).
   One normally uses $\beta = 1$. \richard{Je crois que ce facteur 
\texttt{beta} devrait dispara\^\i tre car des poids beaucoup plus
g\'en\'eraux sont maintenant implant\'es dans les classes  \texttt{*Weights}.}
  \endtabb
\code

   virtual ~Normalizer () \hide { delete [] m_cst; } \endhide 
\endcode
 \tabb 
  Destructor.
 \endtabb
\code

   void init (const MScal & m, int k, double beta);
\endcode
 \tabb 
   Initializes the bounds on the length of the shortest vector. The lattices
   have $m$ points per unit volume, are of rank $k$, and the bias factor is 
   \texttt{beta} for all dimensions $j \le $ \texttt{maxDim}.
 \endtabb
\code

   std::string ToString () const;
\endcode
 \tabb 
   Returns this object as a string.
 \endtabb
\code

   NormType getNorm () const \hide { return m_norm; } \endhide
\endcode
 \tabb 
   Returns the norm associated with this object.
 \endtabb
\code

   void setNorm (NormType norm) \hide { m_norm = norm; } \endhide
\endcode
 \tabb 
   Sets the norm associated with this object to \texttt{norm}.
 \endtabb
\code

   int getDim () const \hide { return m_maxDim; } \endhide 
\endcode
 \tabb 
   Returns the maximal dimension for this object.
 \endtabb
\code

   double & getCst (int j);
\endcode
 \tabb 
   Returns the bound on the length of the shortest nonzero vector in
  dimension $j$.
% Indices $j$ goes from 1 to the value returned by \texttt{dim}.
 \endtabb
\code

   virtual double getGamma (int j) const;
\endcode
 \tabb 
   Returns the value of the lattice constant $\gamma_j$ in dimension  $j$.
  For this base class, always returns 1.
 \endtabb
\code


protected:

   std::string m_name;
\endcode
 \tabb 
  Name of the normalizer.
 \endtabb
\code

   NormType m_norm;
\endcode
 \tabb 
  Norm associated with this object.
 \endtabb
\code

   MScal m_m;
\endcode
 \tabb 
  Number of points of the lattice per unit volume.
 \endtabb
\code

   int m_rank;
\endcode
 \tabb 
  Rank of the lattice.
 \endtabb
\code

   int m_maxDim;
\endcode
 \tabb 
  Only elements 1 to \texttt{m\_maxDim} (inclusive) of arrays are defined.
 \endtabb
\code

   double m_beta;
\endcode
 \tabb 
  Beta factor.
 \endtabb
\code

   double *m_cst;
\endcode
 \tabb 
   Contains the bounds on the length of the shortest nonzero
    vector in the lattice in each dimension.
 \endtabb
\ifdetailed
\code


private:

   Normalizer (const Normalizer &);
   Normalizer & operator= (const Normalizer &);
\endcode
 \tabb 
   Use of the copy-constructor and of assigment are forbidden.
 \endtabb
\fi
\code

};

}

\hide
#endif
\endhide
\endcode
