% Copyright (c) 2012 Richard Simard, Pierre L'Ecuyer, Université de Montréal.
% 
% This file is part of Lattice Builder.
% 
% Lattice Builder is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
% 
% Lattice Builder is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
% 
% You should have received a copy of the GNU General Public License
% along with Lattice Builder.  If not, see <http://www.gnu.org/licenses/>.

\defmodule {Random}

This class generates random numbers (in fact pseudo-random numbers).
The generator used is the 64-bits generator \texttt{LFSR258}
from L'Ecuyer \cite{rLEC99a}  with period length near $2^{258}$
for 64-bits machines, and the 32-bits generator \texttt{LFSR113}
from L'Ecuyer \cite{rLEC99a}  with period length near $2^{113}$
on  32-bits machines. Thus the random numbers generated will be
different on 32-bits and 64-bits machines.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule
\code\hide
#ifndef RANDOM_H
#define RANDOM_H
\endhide

namespace LatCommon {

class Random {
public:

   Random();
\endcode
  \tabb Constructor using a default seed for the generator.
   One may reset the seed by calling \texttt{setSeed}.
  \endtabb
\code

   ~Random() \hide {} \endhide
\endcode
\tabb
Destructor.
\endtabb
\code

   double randU01();
\endcode
\tabb
Returns a random number in $[0, 1)$. The number has 53 random bits
of resolution on 64-bits machines, and 32 random bits 
on 32-bits machines.
\endtabb
\code

   int randInt (int i, int j);
\endcode
\tabb
Return a random integer in $[i, j]$. The numbers $i$ and $j$ can occur.
Restriction: $i < j$.
\endtabb
\code

   unsigned long randBits (int s);
\endcode
\tabb Returns random blocks of $s$ bits ($s$-bit integers).
\endtabb
\code

   void setSeed (unsigned long seed);
\endcode
\tabb
Sets the seed of the generator. If not called, a default seed is used.
\endtabb
\code


private:

   unsigned long randValue();

   unsigned long etat1, etat2, etat3, etat4, etat5;
};

}
\hide
#endif
\endhide
\endcode
