% Copyright (c) 2012 Richard Simard, Pierre L'Ecuyer, Université de Montréal.
% 
% This file is part of Lattice Builder.
% 
% Lattice Builder is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
% 
% Lattice Builder is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
% 
% You should have received a copy of the GNU General Public License
% along with Lattice Builder.  If not, see <http://www.gnu.org/licenses/>.

\defmodule {Rank1Lattice}
\hpierre{Il faut pouvoir choisir chaque $a_i$ ind\'ependamment}

This class implements a general rank 1 lattice basis. For the values
 $a_1, a_2, \ldots, a_{d}$ given, the $d$-dimensional lattice basis is formed
 as:
\[
\mathbf{b_1} = (a_1, a_2, \ldots, a_{d}),\quad \mathbf{b_2} = (0, n, 0, \ldots, 0),\quad  \ldots, \quad \mathbf{b_d} = (0, \ldots, 0, n)
\]
Without loss of generality, one may choose $a_1 = 1$.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule
\code\hide
#ifndef RANK1LATTICE_H
#define RANK1LATTICE_H
\endhide
#include "latcommon/Types.h"
#include "latcommon/Const.h"
#include "latcommon/IntLattice.h"


namespace LatCommon {

class Rank1Lattice: public IntLattice {
public:

   Rank1Lattice (const MScal & n, const MVect & a, int d,
                 NormType norm = L2NORM);
\endcode
\tabb
Constructor. $d$ represents the number of multipliers in the array \texttt{a}.
\endtabb
\code

   Rank1Lattice (const Rank1Lattice & Lat);
\endcode
\tabb
  Copy constructor.
\endtabb
\code

   Rank1Lattice & operator= (const Rank1Lattice & Lat);
\endcode
\tabb
 Assigns \texttt{Lat} to this object.
\endtabb
\code

   ~Rank1Lattice();
\endcode
\tabb
Destructor.
\endtabb
\code

   std::string toStringCoef() const;
\endcode
\tabb
Returns the vector of multipliers $a$ as a string.
\endtabb
\code

   void buildBasis (int d);
\endcode
\tabb
Builds the basis in dimension $d$.
\endtabb
\code

   void incDim ();
\endcode
\tabb
Increases the dimension by 1.
\endtabb
\code


protected:

   void init();
\endcode
\tabb
Initializes the rank 1 lattice.
\endtabb
\code

   MVect m_a;
\endcode
\tabb
The multipliers of the rank 1 lattice rule.
\endtabb
\code
};

}
\hide
#endif
\endhide
\endcode
